---
title: "Untitled"
output:
  revealjs::revealjs_presentation:
    theme: default
    highlight: default
    center: false
    transition: fade
    background_transition: none
    self_contained: no
    css: custom.css
---

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide header

Content shown first

. . .

Content shown next on the same slide

## regular header at top

This is a slide
 
## fixed header at top {.special-header}

:::: {.special-body}
This is a slide
::::

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE, out.extra='class = "plain"'}
plot(cars)
```

